---
title: "Morris Screening"
author: "Jason Hilton"
date: "23 May 2019"
output: html_document
---


Define a funky test function

$$
y= \beta_0 + \beta_1x_1 + \beta_2x_1x_2 + \beta_4x_1^4 + \beta_5\text{sin}(4x_2\pi)
$$


```{r}
library(tibble)
library(ggplot2)
library(magrittr)
library(DiceKriging)
library(sensitivity)
library(lhs)
library(dplyr)

set.seed(1434)
test_function <- function(X, B){
  cbind(1, X, X[,1] * X[,2], X[,1]*X[,1]*X[,1]*X[,1], 
        sin(X[,2]*4*pi)) %*% as.matrix(B)
}

# define the true function so that it ignores any additional columns of X
get_true_function <- function(true_beta){
  true_func <- function(X){
    test_function(X[,1:2], true_beta)
  }
  return(true_func)
}

# some arbitrary coefficients
true_beta <- c(3,-0.5,0.25, 2, -1/10,1)
true_function  <- get_true_function(true_beta)

```


What does this look like

```{r}


k <- 2
N <- 100
x <- seq(0,1, 0.05)

# generate a simple grid like design for ease of plotting
X <- as.matrix(expand.grid(replicate(k, x, simplify = F)))
# Observations
Y <- true_function(X)

# stick this in a single dataset
X_df <- as_tibble(X) %>% mutate(y=Y)
X_df
```



```{r}

ggplot(X_df,aes(x=Var1,y=y, colour=Var2)) + geom_line()+ scale_colour_viridis_c() +
  facet_wrap(~Var2)
ggplot(X_df,aes(x=Var2,y=y, colour=Var1)) + geom_line()+ scale_colour_viridis_c() +
  facet_wrap(~Var1)
ggplot(X_df, aes(x=Var1, y=Var2, fill=y)) + geom_tile() + scale_fill_viridis_c()

```


# Morris method

This method picks somepoints in design space and generates very simple paths through parameter space starting from each point, varying one input at a time (One-At-a-Time (OAT)).
The mean effect on the output of the variation in each direction and the variance of these (accross sample starting points) describe the linear and non-linear effects of each variable.

Pretend we have 20 inputs (all but the first two are inactive), and we try r=10 starting points.
This needs 20*10= 200 simulation runs points. It's a slightly odd looking design.

```{r}
# is the number of points
D1 <- morris(factors=20, r=10, 
       design=list(type="oat",
                   levels=6,grid.jump=2))


plot(D1$X[,1], D1$X[,2]) #

```

Generate the outputs from these points. All linear and not linear effects are zero, except the ones we want to keep.

```{r}
Y1 <- true_function(D1$X)
tell(D1, Y1) # Modifies the morris object D1
plot(D1) 

print(D1)


```


What if we add noise? If we add too much we obviously won't be able to distinguish the signal.
Let's add one fifth the real signal. This seems to be ok. There are some distributional arguments about the threshold level of sigma.

```{r}

noise_sig <- sd(Y1)/5

D2 <- morris(factors=20, r=25, 
       design=list(type="oat",
                   levels=6,
                   grid.jump=2))

nobs <- dim(D2$X)[1]

Y2 <- true_function(D2$X) + rnorm(nobs, 0, noise_sig)
tell(D2, Y2) # Modifies the morris object D2
plot(D2) 

print(D2)


```

How does the kriging model do in the screened and non-screened spaces? 
Fit the models using the DiceKriging packages.

```{r}
# build an lhs sample in the higher and lower dimensional spaces
D1<-lhs::maximinLHS(n=400, k=2, method = "build", dup=4) 
D2<-lhs::maximinLHS(n=400, k=10, method = "build", dup=4)
D3<-lhs::maximinLHS(n=400, k=20, method = "build", dup=4)
Y1 <- true_function(D1)
nug_var <- sd(Y1)/5
Y1 <- Y1 + rnorm(length(Y1), 0, sd(Y1)/5)
Y2 <- true_function(D2) + rnorm(length(Y1), 0, sd(Y1)/5)
Y3 <- true_function(D3) + rnorm(length(Y1), 0, sd(Y1)/5)

D1_df <- as_tibble(D1)
D2_df <- as_tibble(D2)
D3_df <- as_tibble(D3)
# constant mean
mod1 <- km(Y~1, D1,Y1, covtype="gauss", nugget.estim = T)


```

```{r}

mod1
plot(mod1)
```



```{r}
# predict using our grid from before
X_df<- as_tibble(X[,1:2])
mod1_pred <- predict(mod1, newdata=X_df, checkNames=F,
                 type="SK") # SK for simple kriging - i.e. no linear trend in the mean function


X_df %<>% mutate(Y=true_function(X),
                 Y_hat=mod1_pred$mean,
                 Y_low=mod1_pred$lower95,
                 Y_hi=mod1_pred$upper95,
                 Error=Y-Y_hat,
                 Abs_error=abs(Error),
                 PC_error= Abs_error/Y)

ggplot(X_df, aes(x=Var1,y=Y_hat))  + 
  facet_wrap(~Var2, scales="free") +
   geom_ribbon(aes(ymin=Y_low,ymax=Y_hi), fill="grey") +
  geom_line() + 
  theme_minimal() +geom_point(aes(y=Y))



ggplot(X_df, aes(x=Var2,y=Y_hat))  + 
  facet_wrap(~Var1, scales="free") +
   geom_ribbon(aes(ymin=Y_low,ymax=Y_hi), fill="grey") +
  geom_line() + 
  theme_minimal() +geom_point(aes(y=Y))


X_df %>% summarise(mean(PC_error))

```

# 10 dimensions

```{r}


# constant mean
mod2 <- km(Y~1, D2,Y2, covtype="gauss", nugget.estim = T)
print(mod2)
plot(mod2)
X_df<- as_tibble(cbind(X[,1:2], matrix(0.5, dim(X)[1], 8)))
mod2_pred <- predict(mod2, newdata=X_df, checkNames=F,
                 type="SK") # SK for simple kriging - i.e. no linear trend in the mean function


X_df %<>% mutate(Y=true_function(X),
                 Y_hat=mod2_pred$mean,
                 Y_low=mod2_pred$lower95,
                 Y_hi=mod2_pred$upper95,
                 Error=Y-Y_hat,
                 Abs_error=abs(Error),
                 PC_error= Abs_error/Y)

ggplot(X_df, aes(x=Var1,y=Y_hat))  + 
  facet_wrap(~Var2, scales="free") +
   geom_ribbon(aes(ymin=Y_low,ymax=Y_hi), fill="grey") +
  geom_line() + 
  theme_minimal() +geom_point(aes(y=Y))



ggplot(X_df, aes(x=Var2,y=Y_hat))  + 
  facet_wrap(~Var1, scales="free") +
   geom_ribbon(aes(ymin=Y_low,ymax=Y_hi), fill="grey") +
  geom_line() + 
  theme_minimal() +geom_point(aes(y=Y))


X_df %>% summarise(mean(PC_error))

```


# 20 Dimensions

```{r}

mod3 <- km(Y~1, D3,Y3, covtype="gauss", nugget.estim = T)
print(mod3)
plot(mod3)


X_df<- as_tibble(cbind(X[,1:2], matrix(0.5, dim(X)[1], 18)))
mod3_pred <- predict(mod3, newdata=X_df, checkNames=F,
                 type="SK") # SK for simple kriging - i.e. no linear trend in the mean function



X_df %<>% mutate(Y=true_function(X),
                 Y_hat=mod3_pred$mean,
                 Y_low=mod3_pred$lower95,
                 Y_hi=mod3_pred$upper95,
                 Error=Y-Y_hat,
                 Abs_error=abs(Error),
                 PC_error= Abs_error/Y)

ggplot(X_df, aes(x=Var1,y=Y_hat))  + 
  facet_wrap(~Var2, scales="free") +
   geom_ribbon(aes(ymin=Y_low,ymax=Y_hi), fill="grey") +
  geom_line() + 
  theme_minimal() +geom_point(aes(y=Y))

ggplot(X_df, aes(x=Var2,y=Y_hat))  + 
  facet_wrap(~Var1, scales="free") +
   geom_ribbon(aes(ymin=Y_low,ymax=Y_hi), fill="grey") +
  geom_line() + 
  theme_minimal() +geom_point(aes(y=Y))

X_df %>% summarise(mean(PC_error))


```


## Automatic Relevance detection

Higher points indicate relevance

```{r}


ard_df <- rbind(tibble(Model="20D",
             ScaleLength=coef(mod3)$range,
             Dimension=1:20),
      tibble(Model="10D",
             ScaleLength=coef(mod2)$range,
             Dimension=1:10),
      tibble(Model="2D",
             ScaleLength=coef(mod1)$range,
             Dimension=1:2))


ggplot(ard_df, aes(x=Dimension, y= 1/ScaleLength,colour=Model,shape=Model))+
  geom_point() + ggtitle("Automatic Relevance Detection") + facet_wrap(~Model)

```

## Nugget estimates
```{r}

nug_df <- tibble(Model=c("2d","10d","20d"),
       Nugget_hat= c(coef(mod1)$nugget,
                     coef(mod2)$nugget,
                     coef(mod3)$nugget),
       True_nugget=nug_var**2)

ggplot(nug_df, aes(x=Model,y=Nugget_hat)) + geom_point() + 
  geom_hline(aes(yintercept=True_nugget))


```

## Sensitivity

Sensitivity analysis based on numerically probing the emulator and calculating Sobol indicies

```{r}

wrap_mod <- function(krig_model){
  predict_krig <- function(X){
    predict(krig_model, newdata=X, checkNames=F,
                 type="SK")$mean
  }
  return(predict_krig)
}


sens_f1<- wrap_mod(mod1)
x <- fast99(model = sens_f1, factors = c("X1","X2"), n = 1000,
            q = "qunif", q.arg = list(min = -pi, max = pi))

x

plot(x)
```


```{r}
sens_f2<- wrap_mod(mod2)
x <- fast99(model = sens_f2, factors = 10, n = 5000,
            q = "qunif", q.arg = list(min = -pi, max = pi))

x

plot(x)


```


### 20 Dimensions

```{r}
sens_f3<- wrap_mod(mod3)
x <- fast99(model = sens_f3, factors = 20, n = 10000,
            q = "qunif", q.arg = list(min = -pi, max = pi))

x

plot(x)

```








